---
output:
  pdf_document:
    fig_height: 3
    latex_engine: pdflatex
    keep_tex: true
    #output: pdf_document
header-includes:
   - \usepackage{helvet}
   - \renewcommand*\familydefault{\sfdefault}
   - \usepackage[T1]{fontenc}
   - \usepackage{xcolor}
   - \usepackage{placeins}
   - \usepackage{float}
   - \usepackage{needspace}
   - \usepackage{titlesec}
   - \titlespacing{\section}{0pt}{12pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
   - \titlespacing{\subsection}{0pt}{12pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
   - \titlespacing{\subsubsection}{0pt}{12pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
   - \usepackage{etoolbox}
   - \preto\section{\needspace{5\baselineskip}}
   - \preto\subsection{\needspace{4\baselineskip}}
   - \preto\subsubsection{\needspace{3\baselineskip}}

geometry: margin=0.5in
---

```{r setup, include=FALSE}
# Global chunk options and required libraries
knitr::opts_chunk$set(
    echo = TRUE,
    fig.pos = "H",  # Hold figure position - prevents figures from floating away from their titles
    out.extra = ""  # Additional LaTeX options for figures
)

# Set knitr hook to prevent page breaks before section titles
knitr::knit_hooks$set(plot = function(x, options) {
  paste0("\\nopagebreak\n", knitr::hook_plot_tex(x, options), "\n\\nopagebreak\n")
})

library(ggplot2)
library(scales)
library(data.table)
library(rmarkdown)
library(dplyr)
library(knitr)
library(kableExtra)
library(htmltools)



args <- commandArgs(trailingOnly = TRUE)
if (length(args) < 22) {
  stop("Usage: Rscript report.R <sample_id> <cramino_stat> <sample_ids_file> <votes_file> <dictionary_file> <logo_file> <copy_number_plot_file> <tumor_copy_number_file> <cnv_filter_file> <cnv_chr9> <cnv_chr7> <mgmt_results_file> <snv_results_file> <structure_variant_file> <terp_html>  <svanna_html> <egfr_plot_file> <idh1_plot_file> <idh2_plot_file> <tertp_plot_file> <tsneplot> <snv_target_genes> <output_pdf_file>")
}


sample_id              <- args[1]
cramino_stat           <- args[2]
sample_ids_file        <- args[3]
votes_file             <- args[4]
dictionary_file        <- args[5]
logo_file              <- args[6]
copy_number_plot_file  <- args[7]
tumor_copy_number_file <- args[8]
cnv_filter_file        <- args[9]
cnv_chr9               <- args[10]
cnv_chr7               <- args[11]
mgmt_results_file      <- args[12]
snv_results_file       <- args[13]
structure_variant_file <- args[14]
terp_html              <- args[15]
#annotsv_html           <- args[16]
svanna_html            <- args[16]
egfr_plot_file         <- args[17]
idh1_plot_file         <- args[18]
idh2_plot_file         <- args[19]
tertp_plot_file        <- args[20]
tsneplot               <- args[21]
snv_target_genes       <- args[22]
output_pdf_file        <- args[23]

# Read in the votes and dictionary files
if (!file.exists(votes_file)) {
  stop("Votes file not found: ", votes_file)
}

# Check if votes file is empty before reading
file_info <- file.info(votes_file)
if (file_info$size == 0) {
  stop("Votes file is empty: ", votes_file)
}

# Try reading and check if it has data
votes_lines <- readLines(votes_file, warn = FALSE)
if (length(votes_lines) <= 1) {
  stop("Votes file contains only headers or is empty: ", votes_file)
}

votes <- read.csv(votes_file, header = TRUE, sep = "\t")
if (nrow(votes) == 0) {
  stop("Votes file has no data rows: ", votes_file)
}

print(votes)
mc <- read.csv(votes_file, header = TRUE, sep = "\t")
if (!file.exists(dictionary_file)) {
  stop("Votes file not found: ", dictionary_file)
}
dict <- read.csv(dictionary_file, header = TRUE, sep = "\t")

# Merge the votes with the dictionary
mc <- votes %>% 
  left_join(dict, by = c("class" = "Methylation.Class.Name.Abbreviated"))

# Summarise scores by methylation class family
mcf <- mc %>%
  group_by(Methylation.Class.Family) %>%
  summarise(score_MCF = sum(score))

# Extract mean coverage from cramino statistics
lines <- readLines(cramino_stat)
mean_coverage_line <- lines[grep("Mean coverage", lines, ignore.case = TRUE)]
if (length(mean_coverage_line) > 0) {
  # Extract the numeric value from the line
  mean_coverage <- gsub(".*Mean coverage[^0-9]*([0-9.]+).*", "\\1", mean_coverage_line[1])
} else {
  mean_coverage <- "N/A"
}

# Extract tumor content
samples_df <- tryCatch({
  read.table(sample_ids_file, header = FALSE, sep = "", stringsAsFactors = FALSE)
}, error = function(e) {
  return(data.frame())
})

if (nrow(samples_df) >= 1 && ncol(samples_df) >= 2) {
  colnames(samples_df) <- c("Sample", "TumorContent")
  matched_row <- samples_df[samples_df$Sample == sample_id, ]
  if (nrow(matched_row) > 0) {
    tumor_percent <- round(as.numeric(matched_row$TumorContent) * 100)
  } else {
    tumor_percent <- "N/A"
  }
} else {
  tumor_percent <- "N/A"
}

```

```{r, out.height= "50px", echo=FALSE, fig.align="right"}
knitr::include_graphics(logo_file)
```

# Executive Summary

## Sample ID: `r args[1]`

The report is based on a mean coverage of **`r mean_coverage`** and the estimated tumor content in sample is the **`r tumor_percent`%**.

### CrossNN Methylation-based Classification

```{r echo=FALSE, results='asis'}
# Create the classification summary table
classification_data <- data.frame(
  Classification = c("Methylation Class Family", "Methylation Class"),
  Result = c(
    ifelse(max(mcf$score_MCF) >= 0.2, 
           as.character(mcf[which.max(mcf$score_MCF), ]$Methylation.Class.Family), 
           "not classifiable"),
    ifelse(mc$score[1] >= 0.2, 
           as.character(mc$Methylation.Class.Name[1]), 
           "not classifiable ")
  ),
  Score = c(
    ifelse(max(mcf$score_MCF) >= 0.2, round(max(mcf$score_MCF), digits = 3), "-"),
    ifelse(mc$score[1] >= 0.2, round(mc$score[1], digits = 3), "-")
  )
)

if (knitr::is_latex_output()) {
  kable(classification_data,
        format = "latex",
        booktabs = TRUE,
        col.names = c("Classification", "Result", "Score"),
        align = "lcc",
        row.names = FALSE) %>%
    kable_styling(latex_options = c("striped", "hold_position"),
                  full_width = FALSE,
                  position = "center") %>%
    column_spec(2, bold = TRUE) %>%
    column_spec(3, bold = TRUE)
} else {
  kable(classification_data,
        format = "html",
        col.names = c("Classification", "Result", "Score"),
        align = "lcc",
        row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE,
                  position = "center") %>%
    column_spec(2, bold = TRUE) %>%
    column_spec(3, bold = TRUE)
}
```

### Full CNV Profile


```{r, out.height = "120px", fig.align='center', echo=FALSE, results='asis'}
#cat("`Genes annotated in the full CNV profile are amplified (Gain) or deleted (Loss) based on QDNAseq results.`")
# Display the copy number plot using the provided file path
knitr::include_graphics(copy_number_plot_file)
```

### SNV Calling and Annotation

```{r comment='', echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(knitr)
library(kableExtra)
library(dplyr)

# Read the SNV annotation results and target genes list
tryCatch({
    # Read the SNV results table
    df_snv_all <- read.csv(snv_results_file, stringsAsFactors = FALSE, sep = "\t", row.names = NULL)
    
    # Read the target genes list
    target_genes <- tryCatch({
        readLines(snv_target_genes)
    }, error = function(e) {
        character(0)  # Return empty if file cannot be read
    })
    
    # Handle empty case
    if (nrow(df_snv_all) == 0 || length(target_genes) == 0) {
        cat("No SNV data available or no target genes specified.\n")
    } else {
        # ---- 1. Remove ExonicFunc.refGene ----
        if ("ExonicFunc.refGene" %in% colnames(df_snv_all)) {
            df_snv_all <- df_snv_all %>% select(-ExonicFunc.refGene)
        }

        # ---- 2. Clean COSMIC100 ----
        if ("COSMIC100" %in% colnames(df_snv_all)) {
            df_snv_all$COSMIC100 <- gsub("ID=", "", df_snv_all$COSMIC100)
        }

        # ---- 3. Merge Depth columns ----
        if (all(c("Depth", "ClairS_Depth") %in% colnames(df_snv_all))) {
            df_snv_all <- df_snv_all %>%
                mutate(Depth = pmax(as.numeric(Depth), as.numeric(ClairS_Depth), na.rm = TRUE)) %>%
                select(-ClairS_Depth)
        }

        # ---- 3.5. Filter variants with Depth >= 10 ----
        if ("Depth" %in% colnames(df_snv_all)) {
            df_snv_all <- df_snv_all %>%
                filter(as.numeric(Depth) >= 10)
        }

        # ---- 3.6. Filter by variant caller consensus (at least 2 callers or pathogenic) ----
        if ("Variant_caller" %in% colnames(df_snv_all)) {
            df_snv_all <- df_snv_all %>%
                rowwise() %>%
                mutate(
                    caller_count = sum(c(
                        grepl("Pileup", Variant_caller, ignore.case = TRUE),
                        grepl("Merged", Variant_caller, ignore.case = TRUE),
                        grepl("ClairS_TO", Variant_caller, ignore.case = TRUE)
                    )),
                    is_pathogenic = if ("CLNSIG" %in% colnames(df_snv_all)) {
                        CLNSIG == "Pathogenic"
                    } else {
                        FALSE
                    }
                ) %>%
                filter(caller_count >= 2 | is_pathogenic) %>%
                select(-caller_count, -is_pathogenic) %>%
                ungroup()
        }

        # ---- 4. Create RefSeq and Nucleotide alteration columns from AAChange.refGene ----
        if ("AAChange.refGene" %in% colnames(df_snv_all)) {
            create_nucleotide_column <- function(text) {
                if (is.na(text) || text == "") return("No")
                entries <- unlist(strsplit(text, ","))
                entry <- entries[1]
                parts <- unlist(strsplit(entry, ":"))
                if (length(parts) >= 4) {
                    c_change <- parts[4]
                    return(c_change)
                } else {
                    return("No")
                }
            }

            df_snv_all$`Nucleotide alteration` <- sapply(df_snv_all$AAChange.refGene, create_nucleotide_column)

            # Process AAChange.refGene for protein changes only
            process_aachange <- function(text) {
                if (is.na(text) || text == "") return("No")
                entries <- unlist(strsplit(text, ","))
                patterns <- sapply(entries, function(x) {
                    match <- regexpr("p\\.[^:]*", x)
                    if (match != -1) substr(x, match, match + attr(match, "match.length") - 1) else NA
                })
                patterns <- patterns[!is.na(patterns)]
                if (length(patterns) == 0) return("No")
                if (length(patterns) == 1) return(patterns[1])
                most_frequent <- names(which.max(table(patterns)))
                return(most_frequent)
            }
            df_snv_all$AAChange.refGene <- sapply(df_snv_all$AAChange.refGene, process_aachange)
        }

        # ---- 4.5. Special handling for TERT promoter mutations ----
        if (all(c("Gene.refGene", "Func.refGene", "Start") %in% colnames(df_snv_all))) {
            df_snv_all <- df_snv_all %>%
                mutate(`Nucleotide alteration` = case_when(
                    Gene.refGene == "TERT" & Func.refGene == "upstream" & Start == "1295113" ~ "C228T",
                    Gene.refGene == "TERT" & Func.refGene == "upstream" & Start == "1295135" ~ "C250T",
                    TRUE ~ `Nucleotide alteration`
                ))
        }

        if (all(c("Gene.refGene", "Chr", "Start", "End") %in% colnames(df_snv))) {
            df_snv <- df_snv %>%
                filter(!(Gene.refGene == "TERT" & Chr == "chr5" & Start == "1295957" & End == "1295957"))

        # ---- 5. Filter for target genes OR pathogenic variants ----
        if ("Gene.refGene" %in% colnames(df_snv_all)) {
            # Filter for variants that meet EITHER condition:
            # 1. Gene is in the target genes list (regardless of CLNSIG status), OR
            # 2. CLNSIG = "Pathogenic" (regardless of whether gene is in target list)
            # Examples:
            #   - Gene in target + Pathogenic → INCLUDED
            #   - Gene in target + not Pathogenic → INCLUDED
            #   - Gene NOT in target + Pathogenic → INCLUDED
            #   - Gene NOT in target + not Pathogenic → EXCLUDED
            
            # Check if CLNSIG column exists and apply appropriate filter
            if ("CLNSIG" %in% colnames(df_snv_all)) {
                # Use | (OR) to select rows where EITHER condition is TRUE
                df_snv_filtered <- df_snv_all %>%
                    rowwise() %>%
                    mutate(
                        in_target = Gene.refGene %in% target_genes,
                        is_pathogenic_exact = CLNSIG == "Pathogenic"
                    ) %>%
                    filter(in_target | is_pathogenic_exact) %>%
                    select(-in_target, -is_pathogenic_exact) %>%
                    ungroup()
            } else {
                # If no CLNSIG column, filter by target genes only
                df_snv_filtered <- df_snv_all %>%
                    filter(Gene.refGene %in% target_genes)
            }
            
            # Check if any variants remain after filtering
            if (nrow(df_snv_filtered) == 0) {
                cat("No SNV detected in target genes or pathogenic variants.\n")
            } else {
                # ---- 6. Select only required columns ----
                # Rename Gene.refGene to Gene Name for display
                df_snv_display <- df_snv_filtered %>%
                    select(Chr, `Gene ` = Gene.refGene, `Nucleotide alteration`, `AA Change` = AAChange.refGene, AF, CLNSIG)

                # Remove row names
                rownames(df_snv_display) <- NULL

                # ---- 7. Display the table ----
                # Find Gene column index for italic formatting
                gene_col_idx <- which(colnames(df_snv_display) == "Gene ")

                if (knitr::is_html_output()) {
                    tbl <- kable(df_snv_display,
                                 format = "html",
                                 table.attr = "style='border: 1px solid black; border-collapse: collapse; width: 80%; margin-left: auto; margin-right: auto;'",
                                 align = "c",
                                 row.names = FALSE,
                                 escape = FALSE)
                    for (i in 1:(ncol(df_snv_display) - 1)) {
                        tbl <- tbl %>% column_spec(i, border_right = "1px solid black")
                    }
                    # Apply italic to Gene column
                    if (length(gene_col_idx) > 0) {
                        tbl <- tbl %>% column_spec(gene_col_idx, italic = TRUE)
                    }
                    tbl %>%
                        kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center")
                } else if (knitr::is_latex_output()) {
                    tbl <- kable(df_snv_display,
                          format = "latex",
                          booktabs = TRUE,
                          align = "c",
                          row.names = FALSE) %>%
                        kable_styling(latex_options = c("repeat_header", "striped", "hold_position", "scale_down"),
                                      full_width = FALSE,
                                      position = "center")
                    # Apply italic to Gene column
                    if (length(gene_col_idx) > 0) {
                        tbl <- tbl %>% column_spec(gene_col_idx, italic = TRUE)
                    }
                    tbl
                } else {
                    kable(df_snv_display, row.names = FALSE)
                }
            }
        } else {
            cat("Gene.refGene column not found in SNV data.\n")
        }
    }
}, error = function(e) {
    cat("Error processing SNV target genes data: ", e$message, "\n")
})

```

### Structural Variants/ Fusion Events

```{r comment='', echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

# Function to check if file is empty or has only header
is_empty_file <- function(df) {
    return(nrow(df) == 0)
}

# Try to read the file and handle potential errors
tryCatch({
    # Read the structure variant file - optimized for tab-separated format
    df_sv <- read.table(structure_variant_file, 
                         header = TRUE, 
                         sep = "\t",
                         stringsAsFactors = FALSE,
                         fill = TRUE,
                         quote = "",
                         comment.char = "",
                         na.strings = c("", "NA", "N/A", "na", "n/a"),
                         check.names = FALSE)
    
    # Define essential columns for structural variants (excluding ID)
    essential_columns <- c("chr", "star", "end", "svtype", "breaking", "Genes")
    
    # Keep only essential columns that exist in the data
    available_essential <- essential_columns[essential_columns %in% colnames(df_sv)]
    df_sv <- df_sv[, available_essential, drop = FALSE]
    
    # Check if the file is empty or has only headers
    if (!is_empty_file(df_sv)) {
        # Add explanatory text for structure variants
        #explanatory_text <- paste("Fusion events detected in the sample. The table shows essential breakpoint information including chromosome, position, structural variant type, and associated genes.")
        #cat(strwrap(explanatory_text, width = 110), sep = "\n\n")

        # Find Genes column index for italic formatting
        genes_col_idx <- which(colnames(df_sv) == "Genes")

        if (knitr::is_html_output()) {
            tbl <- kable(df_sv,
                        format = "html",
                        table.attr = "style='border: 1px solid black; border-collapse: collapse; width: 80%; margin-left: auto; margin-right: auto;'",
                        align = "c",
                        row.names = FALSE,
                        escape = FALSE)
            for (i in 1:(ncol(df_sv) - 1)) {
                tbl <- tbl %>% column_spec(i, border_right = "1px solid black")
            }
            # Apply italic to Genes column
            if (length(genes_col_idx) > 0) {
                tbl <- tbl %>% column_spec(genes_col_idx, italic = TRUE)
            }
            tbl %>%
                kable_styling(bootstrap_options = c("striped", "hover"),
                            full_width = FALSE,
                            position = "center")
        } else if (knitr::is_latex_output()) {
            tbl <- kable(df_sv,
                  format = "latex",
                  booktabs = TRUE,
                  align = "c",
                  row.names = FALSE) %>%
                kable_styling(latex_options = c("repeat_header", "striped", "hold_position", "scale_down"),
                            full_width = FALSE,
                            position = "center")
            # Apply italic to Genes column
            if (length(genes_col_idx) > 0) {
                tbl <- tbl %>% column_spec(genes_col_idx, italic = TRUE)
            }
            tbl
        } else {
            kable(df_sv, row.names = FALSE)
        }
    } else {
        # Handle empty file or header-only case
        if (knitr::is_html_output()) {
            cat("<p style='text-align: center; color: #666;'><em>No fusion event was detected in this sample.</em></p>")
        } else if (knitr::is_latex_output()) {
            cat("No fusion event was detected in this sample.")
        } else {
            cat("No fusion event was detected in this sample.")
        }
    }
}, error = function(e) {
    # Handle file reading errors with more detailed information
    error_msg <- paste("Error reading structure variant file:", e$message)
    cat(error_msg, "\n")
    
    if (knitr::is_html_output()) {
        cat("<p style='text-align: center; color: #666;'><em>No fusion event was detected in this sample.</em></p>")
    } else if (knitr::is_latex_output()) {
        cat("No fusion event was detected in this sample.")
    } else {
        cat("No fusion event was detected in this sample.")
    }
})
```

### *MGMT* methylation status

```{r comment='', echo=FALSE, results='asis'}
tryCatch({
    df_mgmt <- read.csv(mgmt_results_file, stringsAsFactors = FALSE)

    if (nrow(df_mgmt) == 0) {
        cat("No MGMT methylation results available for this sample.\n")
    } else {
        # Extract the mean methylation value (CpG 1-98)
        mean_meth_full <- NA
        classification_full <- NA
        
        if ("mean_methylation_full" %in% colnames(df_mgmt)) {
            mean_meth_full <- sprintf("%.2f", as.numeric(df_mgmt$mean_methylation_full[1]))
        }
        
        if ("Classification_by_Full" %in% colnames(df_mgmt)) {
            classification_full <- df_mgmt$Classification_by_Full[1]
        }
        
        # Display the text
        if (!is.na(mean_meth_full) && !is.na(classification_full)) {
            cat(sprintf("The mean methylation is **%s%%**. The methylation status of the *MGMT* promoter region is classified as **%s**.\n", 
                       mean_meth_full, classification_full))
        } else {
            cat("MGMT methylation data incomplete.\n")
        }
    }
}, error = function(e) {
    cat("Error reading MGMT results file.\n")
})
```

\newpage

# Nanopore Extended WGS Report

## Sample ID: `r args[1]`

##### Read statistics 

```{r comment='', echo=FALSE}
# Read the cramino statistics file and extract specific information
lines <- readLines(cramino_stat)

# Define the information we want to extract
info_to_extract <- c(
  "Number of alignments",
  "% from total alignments", 
  "Number of reads",
  "Mean coverage"
)

# Extract and display each piece of information
for (info in info_to_extract) {
  matching_lines <- lines[grep(info, lines, ignore.case = TRUE)]
  if (length(matching_lines) > 0) {
    cat(matching_lines, sep = '\n')
  } else {
    cat(paste(info, "information not found in the file.\n"))
  }
}

```

```{r comment='', echo=FALSE}
# Get the sample ID from command-line-style arguments
args <- commandArgs(trailingOnly = TRUE)
sample_id <- args[1]

# Read the file using general whitespace separator
samples_df <- tryCatch({
  read.table(args[3], header = FALSE, sep = "", stringsAsFactors = FALSE)
}, error = function(e) {
  return(data.frame())  # Return empty if file cannot be read
})


# Check if file has content
if (nrow(samples_df) == 0) {
  cat("No tumor content provided.\n")
} else {
  # Check if we have the expected number of columns before assigning names
  if (ncol(samples_df) >= 2) {
    colnames(samples_df) <- c("Sample", "TumorContent")
    
    # Match the sample ID exactly
    matched_row <- samples_df[samples_df$Sample == sample_id, ]

    if (nrow(matched_row) == 0) {
      cat(paste0("Sample ID '", sample_id, "' not found in tumor content file.\n"))
    } else {
      tumor_percent <- round(as.numeric(matched_row$TumorContent) * 100)
      cat(paste0("The estimated tumor cell content for ", sample_id, " is: ", tumor_percent, "%\n"))
    }
  } else {
    cat("Estimated tumor content file has insufficient columns. Expected at least 2 columns.\n")
  }
}


```
### CrossNN Methylation-based Classification

Methylation-based classification is based on **`r as.integer(votes$num_features[1])`** CpG sites (overlap of sites covered in this sample and the model).
At the methylation class (MC) level, the sample has been classified as **`r ifelse(mc$score[1] >= 0.2, mc$Methylation.Class.Name[1], "not classifiable (score < 0.2)")`**. 
This prediction has a confidence score of **`r round(mc$score[1], digits=3)`**.
At the methylation class **family** (MCF) level, the sample has been classified as **`r ifelse(max(mcf$score_MCF)>=0.2, mcf[which.max(mcf$score_MCF),]$Methylation.Class.Family, "not classifiable (score < 0.2)")`**.
The MCF prediction has a confidence score of **`r round(max(mcf$score_MCF), digits=3)`**.

Scores for the Top 5 entities on MC and MCF level are given below.
Vertical dashed lines indicate the recommended >0.2 cut-off for classification.

```{r, out.height = "150px", fig.align = 'center', echo=FALSE, warning=FALSE}
library(ggplot2)

# Filter out rows with NA or zero scores before plotting
mc_filtered <- mc %>%
    filter(!is.na(score), score > 0) %>%
    arrange(desc(score)) %>%
    slice_head(n = 5)  # Take top 5 scores

# Only plot if there's data
if (nrow(mc_filtered) > 0) {
    ggplot(data = mc_filtered, aes(x = Methylation.Class.Name, y = score)) +
        geom_bar(stat = "identity", fill = "#1f78b4") +
        geom_hline(yintercept = 0.2, linetype = "dashed") +
        coord_flip() +
        scale_x_discrete(limits = rev(mc_filtered$Methylation.Class.Name)) +
        scale_y_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1), limits = c(0, 1)) +
        ylab("score") + xlab("") +
        ggtitle("Methylation class (MC)") +
        theme_classic() + theme(aspect.ratio = 1)
} else {
    cat("No methylation class scores available for plotting.\n")
}

```

```{r, out.height = "200px", fig.align = 'center', echo=FALSE, warning=FALSE}
# Filter and prepare MCF data
mcf_filtered <- mcf %>%
    filter(!is.na(score_MCF), score_MCF > 0) %>%
    arrange(desc(score_MCF)) %>%
    slice_head(n = 5)  # Take top 5 scores

# Only plot if there's data
if (nrow(mcf_filtered) > 0) {
    ggplot(mcf_filtered, aes(x = Methylation.Class.Family, y = score_MCF)) +
        geom_bar(stat = "identity", fill = "#4daf4a") +
        geom_hline(yintercept = 0.2, linetype = "dashed") +
        coord_flip() +
        scale_x_discrete(limits = rev(mcf_filtered$Methylation.Class.Family)) +
        scale_y_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1), limits = c(0, 1)) +
        ylab("score") + xlab("") +
        ggtitle("Methylation class family (MCF)") +
        theme_classic() + theme(aspect.ratio = 1)
} else {
    cat("No methylation class family scores available for plotting.\n")
}

```
\newpage
## Dimensionality Reduction Plot

The dimensionality reduction plot are only intended for visual inspection, not classification.  
```{r, out.height = "250px", fig.align = 'center', echo=FALSE}
# Include t-SNE plot for methylation classification
knitr::include_graphics(tsneplot)
```


## Copy Number Variation Profile

### Full CNV Profile


```{r, out.height = "150px", echo=FALSE, results='asis'}
cat("`Genes annotated in the full CNV profile are amplified (Gain) or deleted (Loss) based on QDNAseq results.`")
# Display the copy number plot using the provided file path
knitr::include_graphics(copy_number_plot_file)
```
```{r comment='', echo=FALSE}
# Read the CNV data
df <- read.csv(tumor_copy_number_file, sep = "\t", stringsAsFactors = FALSE)

# Define chromosome arms coordinates (GRCh38/hg38) - Autosomes only (1-22)
chr_arms <- data.frame(
    Chromosome = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", 
                   "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", 
                   "21", "22"),
    p_arm_start = rep(1, 22),
    p_arm_end = c(121700000, 91800000, 87800000, 48200000, 46100000, 58500000, 
                  58100000, 43200000, 42200000, 38000000, 51000000, 33200000, 
                  16500000, 16100000, 17083673, 35300000, 22200000, 15400000, 
                  24200000, 26100000, 11000000, 13700000),
    q_arm_start = c(124535434, 96509374, 91711528, 51743951, 50059807, 59128983, 
                    61054331, 45877265, 49632404, 41277936, 53425074, 35151254, 
                    17958183, 17578454, 19725254, 36753077, 25263006, 18460898, 
                    26369569, 28510720, 12954789, 15054318),
    q_arm_end = c(248956422, 242193529, 198295559, 190214555, 181538259, 170805979, 
                  159345973, 145138636, 138394717, 133797422, 135086622, 133275309, 
                  114364328, 107043718, 101991189, 90338345, 83257441, 80373285, 
                  58617616, 64444167, 46709983, 50818468),
    stringsAsFactors = FALSE
)

# Function to check chromosome arm events
check_arm_events <- function(df, chrom, arm_type, start_pos, end_pos, threshold = 0.90) {
    # Get segments for the chromosome arm
    segments <- df[df$Chrom == as.character(chrom), ]
    
    if (arm_type == "p") {
        segments <- segments[segments$End <= end_pos, ]
    } else if (arm_type == "q") {
        segments <- segments[segments$Start >= start_pos, ]
    }
    
    # Check if there are any segments
    if (nrow(segments) == 0) {
        return(list(
            gain_pass = FALSE, 
            loss_pass = FALSE, 
            gain_pct = 0, 
            loss_pct = 0, 
            gain_count = 0,
            loss_count = 0,
            total = 0
        ))
    }
    
    # Count events - handle NA values
    gain_count <- sum(segments[,"Event.Type"] == "Gain", na.rm = TRUE)
    loss_count <- sum(segments[,"Event.Type"] == "Loss", na.rm = TRUE)
    total_count <- nrow(segments)
    
    # Calculate percentages - ensure no division by zero
    gain_pct <- if (total_count > 0) gain_count / total_count else 0
    loss_pct <- if (total_count > 0) loss_count / total_count else 0
    
    # Ensure boolean values (not NA)
    gain_pass <- !is.na(gain_pct) && gain_pct >= threshold
    loss_pass <- !is.na(loss_pct) && loss_pct >= threshold
    
    return(list(
        gain_pass = gain_pass,
        loss_pass = loss_pass,
        gain_pct = gain_pct,
        loss_pct = loss_pct,
        gain_count = gain_count,
        loss_count = loss_count,
        total = total_count
    ))
}

# Check all chromosome arms
results <- list()
for (i in 1:nrow(chr_arms)) {
    chrom <- chr_arms$Chromosome[i]
    
    # Check p-arm
    p_result <- check_arm_events(df, chrom, "p", 
                                  chr_arms$p_arm_start[i], 
                                  chr_arms$p_arm_end[i])
    
    # Check q-arm
    q_result <- check_arm_events(df, chrom, "q", 
                                  chr_arms$q_arm_start[i], 
                                  chr_arms$q_arm_end[i])
    
    # Check if both arms individually meet the threshold for the same event type
    if (isTRUE(p_result$gain_pass) && isTRUE(q_result$gain_pass) && 
        p_result$gain_pct >= 0.90 && q_result$gain_pct >= 0.90) {
        # Both arms have ≥90% gain individually - report whole chromosome gain
        results[[length(results) + 1]] <- sprintf("Chr%s: Gain", chrom)
    } else if (isTRUE(p_result$loss_pass) && isTRUE(q_result$loss_pass) && 
               p_result$loss_pct >= 0.90 && q_result$loss_pct >= 0.90) {
        # Both arms have ≥90% loss individually - report whole chromosome loss
        results[[length(results) + 1]] <- sprintf("Chr%s: Loss", chrom)
    } else {
        # Report individual arms only if whole chromosome event doesn't apply
        # For chromosomes 13, 14, 15, 21, and 22, don't include arm designation
        if (chrom %in% c("13", "14", "15", "21", "22")) {
            if (isTRUE(p_result$gain_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%s: Gain", chrom)
            }
            if (isTRUE(p_result$loss_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%s: Loss", chrom)
            }
            
            if (isTRUE(q_result$gain_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%s: Gain", chrom)
            }
            if (isTRUE(q_result$loss_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%s: Loss", chrom)
            }
        } else {
            # For other chromosomes, include arm designation
            if (isTRUE(p_result$gain_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%sp: Gain", chrom)
            }
            if (isTRUE(p_result$loss_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%sp: Loss", chrom)
            }
            
            if (isTRUE(q_result$gain_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%sq: Gain", chrom)
            }
            if (isTRUE(q_result$loss_pass)) {
                results[[length(results) + 1]] <- sprintf("Chr%sq: Loss", chrom)
            }
        }
    }
}

# Display results
if (length(results) > 0) {
    cat("Chromosome arm-level alterations detected (>90% threshold):\n\n")
    for (result in results) {
        cat("- ", result, "\n")
    }
} else {
    cat("No significant chromosome arm-level alterations detected (>90% threshold).\n")
}
```

### High-Confidence Copy Number Variation in Cancer Genes (Amplifications & Homozygous Deletions)
```{r comment='', echo=FALSE}

#cat("Table 1: Copy Number Variation Filter Table\n\n")
# Read and process the CNV data
# Read and process the CNV data
df_mgmt <- read.csv(cnv_filter_file, stringsAsFactors = FALSE, sep = ',')

# Filter out rows with chrX and convert Score to numeric
df_mgmt$ScoreValue <- as.numeric(sub("SCORE=", "", df_mgmt$Score))

# Detect chromosome column (supports multiple naming conventions)
chrom_col <- grep("CHROM|Chrom|chr", colnames(df_mgmt), value = TRUE)[1]
if (is.na(chrom_col)) {
    stop("Could not find chromosome column. Available columns: ", 
         paste(colnames(df_mgmt), collapse = ", "))
}

# Clean up prefixes
df_mgmt$End <- gsub("END=", "", df_mgmt$End)
df_mgmt$SVLEN <- gsub("SVLEN=", "", df_mgmt$SVLEN)
df_mgmt$Score <- gsub("SCORE=", "", df_mgmt$Score)
df_mgmt$LOG2CNT <- gsub("LOG2CNT=", "", df_mgmt$LOG2CNT)
df_mgmt$Gene <- gsub("gene_name=", "", df_mgmt$Gene)

# Apply filtering: score and autosomes only
df_filtered <- subset(df_mgmt, ScoreValue %in% c(2, -2) & !grepl("X", get(chrom_col)))
df_filtered$ScoreValue <- NULL  # Remove helper column

# Display content
if(nrow(df_filtered) == 0){
    cat("No copy number variation to be reported.\n")
} else {
    # Show explanatory text before the table
    #cat("The table is filtered for copy number variation events with a score of 2 (amplification) or -2 (homozygous deletion) and no sex chromosome.\n\n")
    cat(strwrap("The table is filtered for copy number variation events with a score of 2 (amplification) or -2 (homozygous deletion) and no sex chromosomes.\n",
  width = 110
), sep = "\n")

    # Find Gene column index for italic formatting
    gene_col_idx <- which(colnames(df_filtered) == "Gene")

    # HTML formatting
    if (knitr::is_html_output()) {
        tbl <- kable(df_filtered,
                     format = "html",
                     table.attr = "style='border: 1px solid black; border-collapse: collapse; width: 80%; margin-left: auto; margin-right: auto;'",
                     align = "c",
                     row.names = FALSE,
                     escape = FALSE)
        for (i in 1:(ncol(df_filtered) - 1)) {
            tbl <- tbl %>% column_spec(i, border_right = "1px solid black")
        }
        # Apply italic to Gene column
        if (length(gene_col_idx) > 0) {
            tbl <- tbl %>% column_spec(gene_col_idx, italic = TRUE)
        }
        tbl %>%
            kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center")

    # LaTeX formatting
    } else if (knitr::is_latex_output()) {
        tbl <- kable(df_filtered,
              format = "latex",
              booktabs = TRUE,
              align = "c",
              row.names = FALSE) %>%
            kable_styling(latex_options = c("repeat_header", "striped", "hold_position", "scale_down"),
                         full_width = FALSE,
                         position = "center")
        # Apply italic to Gene column
        if (length(gene_col_idx) > 0) {
            tbl <- tbl %>% column_spec(gene_col_idx, italic = TRUE)
        }
        tbl

    # Fallback for other formats
    } else {
        kable(df_filtered, row.names = FALSE)
    }
}

```

```{r check_egfr_presence, echo=FALSE}
# Check if EGFR gene is present in the CNV table
has_egfr <- FALSE
if (exists("df_filtered") && is.data.frame(df_filtered) && nrow(df_filtered) > 0 && "Gene" %in% colnames(df_filtered)) {
    has_egfr <- "EGFR" %in% df_filtered$Gene
}
```

```{r egfr_section_header, echo=FALSE, eval=has_egfr, results='asis'}
cat("\n\n### Chromosome 7: *EGFR* Profile\n\n")
cat("Chromosome 7 CNV for visual inspection of *EGFR* annotated.\n\n")
```

```{r egfr_cnv_plot, echo=FALSE, eval=has_egfr, out.height="150px"}
# Display the copy number plot using the provided file path
knitr::include_graphics(cnv_chr7)
```

```{r egfr_copynumber_text, echo=FALSE, eval=has_egfr, comment='', message=FALSE, warning=FALSE, results='asis'}
# Read the tumor copy number file
df <- read.csv(tumor_copy_number_file, sep="\t", stringsAsFactors=FALSE)

# Filter for chromosome 7 and get max copy number
chr7_max_cn <- df[df$Chrom == "7", "Tumor.Copy"]
if (length(chr7_max_cn) > 0) {
    max_value <- max(chr7_max_cn, na.rm=TRUE)

    # Format the output with bold text
     if (knitr::is_latex_output()) {
        cat(sprintf("The *EGFR* copy number for the sample %s is **%.2f**", sample_id, max_value))
    } else if (knitr::is_html_output()) {
        cat(sprintf("The *EGFR* copy number for the sample %s is **%.2f**", sample_id, max_value))
    } else {
        cat(sprintf("The *EGFR* copy number for the sample %s is **%.2f**", sample_id, max_value))
    }
} else {
    cat(sprintf("No copy number data found for chromosome 7 in the sample %s", sample_id))
}
cat(". The vertical red line highlights exons 2-7. Deletion of exons 2-7 results in *EGFR*viii variant.\n\n")
```

```{r egfr_exon_plot, echo=FALSE, eval=has_egfr, out.width="0.5\\linewidth", out.height="0.3\\textheight", fig.align='center'}
knitr::include_graphics(egfr_plot_file)
```

### Chromosome 9: *CDKN2A/B* Profile


Chromosome 9 CNV for visual inspection of *CDKN2A/B* annotated.

```{r, out.height = "150px", echo=FALSE}
# Display the copy number plot using the provided file path
knitr::include_graphics(cnv_chr9)
```

```{r comment='', echo=FALSE, message=FALSE, warning=FALSE, results='markup'}
library(knitr)

# Read the copy number file
df <- read.csv(tumor_copy_number_file,
               sep = "\t",
               stringsAsFactors = FALSE)

# Check for CDKN2A/B deletions and their types
cdk2na_hemi <- any(df$Chrom == "9" & 
                   df$Start == 21950000 & 
                   df[,"Event.Type"] == "Loss" &
                   df[,"Change.Value"] == 1)

cdk2na_homo <- any(df$Chrom == "9" & 
                   df$Start == 21950000 & 
                   df[,"Event.Type"] == "Loss" &
                   df[,"Change.Value"] == 2)

cdk2nb_hemi <- any(df$Chrom == "9" & 
                   df$Start == 22000000 & 
                   df[,"Event.Type"] == "Loss" &
                   df[,"Change.Value"] == 1)

cdk2nb_homo <- any(df$Chrom == "9" & 
                   df$Start == 22000000 & 
                   df[,"Event.Type"] == "Loss" &
                   df[,"Change.Value"] == 2)

# Generate and display CDKN2A/B status messages
if (cdk2na_hemi) {
    knitr::asis_output("There is a hemizygous deletion in chromosome 9: *CDKN2A* is deleted\n")
}
if (cdk2na_homo) {
    knitr::asis_output("There is a homozygous deletion in chromosome 9: *CDKN2A* is deleted\n")
}
if (cdk2nb_hemi) {
    knitr::asis_output("There is a hemizygous deletion in chromosome 9: *CDKN2B* is deleted\n")
}
if (cdk2nb_homo) {
    knitr::asis_output("There is a homozygous deletion in chromosome 9: *CDKN2B* is deleted\n")
}
if (!cdk2na_hemi && !cdk2na_homo && !cdk2nb_hemi && !cdk2nb_homo) {
    knitr::asis_output("No *CDKN2A/B* deletions detected\n")
}
```

### *IDH1* p.R132 and *IDH2* p.R172


Sequenced coverage of *IDH1* p.R132 and *IDH2* p.R172 (red horizontal line) for visual inspection.

```{r, out.width="0.49\\linewidth", out.height="0.3\\textheight", fig.show='hold', echo=FALSE}
knitr::include_graphics(c(idh1_plot_file, idh2_plot_file))
```

### *TERT*p

Sequenced coverage of *TERT*p C228T (left vertical red line) and C250T (right vertical red line) for visual inspection.

```{r, out.width="0.5\\linewidth", out.height="0.3\\textheight", fig.align='center', echo=FALSE}
knitr::include_graphics(tertp_plot_file)
```


### *MGMT* methylation status

```{r comment='', echo=FALSE, results='asis'}
tryCatch({
    df_mgmt <- read.csv(mgmt_results_file, stringsAsFactors = FALSE)

    if (nrow(df_mgmt) == 0) {
        cat("No *MGMT* methylation results available for this sample.\n")
    } else {
        cat("Nanopore sequencing covers the methylation status on all 98 CpG sites in the *MGMT* promoter region as opposed to\n")
        cat("the 4 CpG sites (sites 76-79) previously reported. The table below summarizes the methylation level on all\n")
        cat("CpG sites in the *MGMT* promoter region (sites 1-98) and the four previously reported sites (sites 76-79)\n")
        cat("and states how the sample classifies according to both methods.\n\n")

        # Continue with normal table processing
        df_mgmt$sample_id <- sample_id
        df_mgmt <- df_mgmt[, c("sample_id", setdiff(names(df_mgmt), "sample_id"))]

        # Convert methylation values to percentages (multiply by 100)
        if ("mean_methylation_full" %in% colnames(df_mgmt)) {
            df_mgmt$mean_methylation_full <- sprintf("%.2f", as.numeric(df_mgmt$mean_methylation_full))
        }
        if ("mean_methylation_pyro" %in% colnames(df_mgmt)) {
            df_mgmt$mean_methylation_pyro <- sprintf("%.2f", as.numeric(df_mgmt$mean_methylation_pyro))
        }

        # Update column names with new headers and percentage indicators
        colnames(df_mgmt) <- gsub("mean_methylation_full", "Mean Methylation, CpG 1–98 [%]", colnames(df_mgmt))
        colnames(df_mgmt) <- gsub("mean_methylation_pyro", "Mean Methylation, CpG 76–79 [%]", colnames(df_mgmt))
        colnames(df_mgmt) <- gsub("Classification_by_Pyro", "Classification by CpG 76–79", colnames(df_mgmt))
        colnames(df_mgmt) <- gsub("Classification_by_Full", "Classification by CpG 1–98", colnames(df_mgmt))

        if (knitr::is_html_output()) {
            tbl <- kable(df_mgmt,
                         format = "html",
                         table.attr = "style='border: 1px solid black; border-collapse: collapse; width: 80%; margin-left: auto; margin-right: auto;'",
                         align = "c",
                         row.names = FALSE,
                         escape = FALSE)
            for (i in 1:(ncol(df_mgmt) - 1)) {
                tbl <- tbl %>% column_spec(i, border_right = "1px solid black")
            }
            tbl %>%
                kable_styling(bootstrap_options = c("striped", "hover"),
                              full_width = FALSE,
                              position = "center")
        } else if (knitr::is_latex_output()) {
            kable(df_mgmt,
                  format = "latex",
                  booktabs = TRUE,
                  align = "c",
                  row.names = FALSE) %>%
                kable_styling(latex_options = c("repeat_header", "striped", "hold_position", "scale_down"),
                              full_width = FALSE,
                              position = "center")
        } else {
            kable(df_mgmt, row.names = FALSE)
        }
    }
}, error = function(e) {
    cat("Error reading MGMT results file.\n")
})


```

### SNV Calling and Annotation

```{r comment='', echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(knitr)
library(kableExtra)
library(dplyr)
# Read the SNV annotation results using the provided file path
tryCatch({
    # Read the SNV results table
    df_snv <- read.csv(snv_results_file, stringsAsFactors = FALSE, sep = "\t", row.names = NULL)

    # Handle empty case
    if (nrow(df_snv) == 0) {
        cat("No SNV Calling and Annotation detected.\n")
    } else {
        # Display explanatory text when SNVs are present
        cat("SNVs in the select genes. Only non-synonymous exonic variants that are not known to be benign according to ClinVar_20240611 and known pathogenic upstream variants, i.e. *TERT*p, are reported. GQ: Genotype Quality, Depth: Sequenced depth, AD: Allele Depth, GT: Allele Genotype, AF: Allele Frequency. The *GRCh38* build reference genome was used for the SNV calling and annotation.\n\n")
        # ---- 1. Remove ExonicFunc.refGene ----
        df_snv <- df_snv %>% select(-ExonicFunc.refGene)

        # ---- 2. Clean COSMIC100 ----
        if ("COSMIC100" %in% colnames(df_snv)) {
            df_snv$COSMIC100 <- gsub("ID=", "", df_snv$COSMIC100)
        }

        # ---- 3. Merge Depth columns ----
        if (all(c("Depth", "ClairS_Depth") %in% colnames(df_snv))) {
            df_snv <- df_snv %>%
                mutate(Depth = pmax(as.numeric(Depth), as.numeric(ClairS_Depth), na.rm = TRUE)) %>%
                select(-ClairS_Depth)
        }

        # ---- 3.5. Filter variants with Depth >= 10 ----
        if ("Depth" %in% colnames(df_snv)) {
            df_snv <- df_snv %>%
                filter(as.numeric(Depth) >= 10)
            
            # Check if any variants remain after filtering
            if (nrow(df_snv) == 0) {
                msg <- "No SNV Calling and Annotation detected with Depth >= 10."
                if (knitr::is_latex_output()) {
                    cat(paste0(msg, "\n"))
                } else {
                    cat(paste0(msg, "\n"))
                }
                return()  # Exit early if no variants meet the depth criteria
            }
        }

        # ---- 3.6. Filter by variant caller consensus (at least 2 callers or pathogenic) ----
        if ("Variant_caller" %in% colnames(df_snv)) {
            df_snv <- df_snv %>%
                rowwise() %>%
                mutate(
                    # Count how many variant callers detected this variant
                    caller_count = sum(c(
                        grepl("Pileup", Variant_caller, ignore.case = TRUE),
                        grepl("Merged", Variant_caller, ignore.case = TRUE),
                        grepl("ClairS_TO", Variant_caller, ignore.case = TRUE)
                    )),
                    # Check if variant is pathogenic in CLNSIG
                    is_pathogenic = if ("CLNSIG" %in% colnames(df_snv)) {
                        grepl("pathogenic", CLNSIG, ignore.case = TRUE)
                    } else {
                        FALSE
                    }
                ) %>%
                # Keep variants if: called by >=2 callers OR is pathogenic
                filter(caller_count >= 2 | is_pathogenic) %>%
                # Remove helper columns
                select(-caller_count, -is_pathogenic) %>%
                ungroup()
            
            # Check if any variants remain after filtering
            if (nrow(df_snv) == 0) {
                msg <- "No SNV Calling and Annotation detected with sufficient variant caller consensus or pathogenic significance."
                if (knitr::is_latex_output()) {
                    cat(paste0(msg, "\n"))
                } else {
                    cat(paste0(msg, "\n"))
                }
                return()  # Exit early if no variants meet the caller consensus criteria
            }
        }

        # ---- 4. Create RefSeq and Nucleotide alteration columns from AAChange.refGene ----
        if ("AAChange.refGene" %in% colnames(df_snv)) {
            # Function to extract RefSeq (NM_ID only)
            create_refseq_column <- function(text) {
                if (is.na(text) || text == "") return("No")
                entries <- unlist(strsplit(text, ","))
                # Process first entry (most common case)
                entry <- entries[1]
                
                # Extract NM identifier
                # Pattern: GENE:NM_XXXXX:exonX:c.CHANGE:p.CHANGE
                parts <- unlist(strsplit(entry, ":"))
                if (length(parts) >= 2) {
                    nm_id <- parts[2]       # NM_XXXXX
                    return(nm_id)
                } else {
                    return("No")
                }
            }
            
            # Function to extract Nucleotide alteration (c.Change only)
            create_nucleotide_column <- function(text) {
                if (is.na(text) || text == "") return("No")
                entries <- unlist(strsplit(text, ","))
                # Process first entry (most common case)
                entry <- entries[1]
                
                # Extract c.DNA change
                # Pattern: GENE:NM_XXXXX:exonX:c.CHANGE:p.CHANGE
                parts <- unlist(strsplit(entry, ":"))
                if (length(parts) >= 4) {
                    c_change <- parts[4]    # c.CHANGE
                    return(c_change)
                } else {
                    return("No")
                }
            }
            
            # Create the RefSeq and Nucleotide alteration columns
            df_snv$RefSeq <- sapply(df_snv$AAChange.refGene, create_refseq_column)
            df_snv$`Nucleotide alteration` <- sapply(df_snv$AAChange.refGene, create_nucleotide_column)
            
            # Process AAChange.refGene for protein changes only (existing functionality)
            process_aachange <- function(text) {
                if (is.na(text) || text == "") return("No")
                entries <- unlist(strsplit(text, ","))
                patterns <- sapply(entries, function(x) {
                    match <- regexpr("p\\.[^:]*", x)
                    if (match != -1) substr(x, match, match + attr(match, "match.length") - 1) else NA
                })
                patterns <- patterns[!is.na(patterns)]
                if (length(patterns) == 0) return("No")
                if (length(patterns) == 1) return(patterns[1])
                most_frequent <- names(which.max(table(patterns)))
                return(most_frequent)
            }
            df_snv$AAChange.refGene <- sapply(df_snv$AAChange.refGene, process_aachange)
        }

        # ---- 4.5. Special handling for TERT promoter mutations ----
        if (all(c("Gene.refGene", "Func.refGene", "Start") %in% colnames(df_snv))) {
            df_snv <- df_snv %>%
                mutate(`Nucleotide alteration` = case_when(
                    Gene.refGene == "TERT" & Func.refGene == "upstream" & Start == "1295113" ~ "C228T",
                    Gene.refGene == "TERT" & Func.refGene == "upstream" & Start == "1295135" ~ "C250T",
                    TRUE ~ `Nucleotide alteration`
                ))
        }

        if (all(c("Gene.refGene", "Chr", "Start", "End") %in% colnames(df_snv))) {
            df_snv <- df_snv %>%
                filter(!(Gene.refGene == "TERT" & Chr == "chr5" & Start == "1295957" & End == "1295957"))
}

        # ---- 5. Reorder columns according to specified order ----
        # Desired order: "RefSeq" "Gene.refGene" "Nucleotide alteration" "Ref" "Alt" "Func.refGene" "AAChange.refGene" "CLNSIG" "COSMIC100" "GQ" "Depth" "AD" "GT" "AF" "Chr" "Start" "End" "Variant_caller"
        desired_order <- c("RefSeq", "Gene.refGene", "Nucleotide alteration", "Ref", "Alt", "Func.refGene", "AAChange.refGene", "CLNSIG", "COSMIC100", "GQ", "Depth", "AD", "GT", "AF", "Chr", "Start", "End", "Variant_caller")
        
        # Get current column names
        current_cols <- names(df_snv)
        
        # Select columns that exist in both desired order and current columns
        available_cols <- desired_order[desired_order %in% current_cols]
        
        # Add any remaining columns that weren't specified in the desired order
        remaining_cols <- current_cols[!current_cols %in% desired_order]
        
        # Combine for final order
        final_order <- c(available_cols, remaining_cols)
        
        # Reorder the dataframe
        df_snv <- df_snv %>% select(all_of(final_order))

        # # ---- 6. Filter out rows with "No" or empty values in RefSeq, Nucleotide alteration, or AAChange.refGene ----
        # if (nrow(df_snv) > 0) {
        #     # Filter out rows where RefSeq, Nucleotide alteration, or AAChange.refGene is "No" or empty
        #     df_snv <- df_snv %>%
        #         filter(
        #             !is.na(RefSeq) & RefSeq != "" & RefSeq != "No",
        #             !is.na(`Nucleotide alteration`) & `Nucleotide alteration` != "" & `Nucleotide alteration` != "No",
        #             !is.na(AAChange.refGene) & AAChange.refGene != "" & AAChange.refGene != "No"
        #         )
        # }

        # Remove row names
        rownames(df_snv) <- NULL

        # ---- 7. Display the table ----
        # Find Gene.refGene column index for italic formatting
        gene_col_idx <- which(colnames(df_snv) == "Gene.refGene")

        if (knitr::is_html_output()) {
            tbl <- kable(df_snv,
                         format = "html",
                         table.attr = "style='border: 1px solid black; border-collapse: collapse; width: 80%; margin-left: auto; margin-right: auto;'",
                         align = "c",
                         row.names = FALSE,
                         escape = FALSE)
            for (i in 1:(ncol(df_snv) - 1)) {
                tbl <- tbl %>% column_spec(i, border_right = "1px solid black")
            }
            # Apply italic to Gene.refGene column
            if (length(gene_col_idx) > 0) {
                tbl <- tbl %>% column_spec(gene_col_idx, italic = TRUE)
            }
            tbl %>%
                kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center")
        } else if (knitr::is_latex_output()) {
            tbl <- kable(df_snv,
                  format = "latex",
                  booktabs = TRUE,
                  align = "c",
                  row.names = FALSE) %>%
                kable_styling(latex_options = c("repeat_header", "striped", "hold_position", "scale_down"),
                              full_width = FALSE,
                              position = "center")
            # Apply italic to Gene.refGene column
            if (length(gene_col_idx) > 0) {
                tbl <- tbl %>% column_spec(gene_col_idx, italic = TRUE)
            }
            tbl
        } else {
            kable(df_snv, row.names = FALSE)
        }
    }
}, error = function(e) {
    msg <- "Error reading SNV results file."
    if (knitr::is_latex_output()) {
        cat(paste0(msg, "\\\\\n"))
    } else {
        cat(paste0(msg, "\n"))
    }
})

```


### Structural Variation/ Fusion Events
```{r comment='', echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

# Function to check if file is empty or has only header
is_empty_file <- function(df) {
    return(nrow(df) == 0)
}

# Try to read the file and handle potential errors
tryCatch({
    # Read the structure variant file - optimized for tab-separated format
    df_mgmt <- read.table(structure_variant_file, 
                         header = TRUE, 
                         sep = "\t",
                         stringsAsFactors = FALSE,
                         fill = TRUE,
                         quote = "",
                         comment.char = "",
                         na.strings = c("", "NA", "N/A", "na", "n/a"),
                         check.names = FALSE)
    
    # Print diagnostic information
    ##cat(sprintf("Successfully read %d rows with %d columns\n", nrow(df_mgmt), ncol(df_mgmt)))
    #cat("Original columns found:", paste(colnames(df_mgmt), collapse = ", "), "\n")
    
    # Define essential columns for structural variants
    # Updated to include new "Features" column that contains exon/intron information
    essential_columns <- c("chr", "star", "end", "ID", "svtype", "breaking", "Genes", "Features")

    # Keep only essential columns that exist in the data
    available_essential <- essential_columns[essential_columns %in% colnames(df_mgmt)]
    df_mgmt <- df_mgmt[, available_essential, drop = FALSE]
    
    # Print filtered column information
    #cat("Essential columns kept:", paste(colnames(df_mgmt), collapse = ", "), "\n")
    
    # Check if the file is empty or has only headers
    if (!is_empty_file(df_mgmt)) {
        # Add explanatory text for structure variants
        explanatory_text <- paste("Fusion events detected in the sample. The table shows essential breakpoint information including chromosome, position, Sniffles ID, structural variant type, associated genes, and genomic features (exons/introns) involved in the fusion.")
        cat(strwrap(explanatory_text, width = 110), sep = "\n\n")

        # Find Genes column index for italic formatting
        genes_col_idx <- which(colnames(df_mgmt) == "Genes")

        if (knitr::is_html_output()) {
            tbl <- kable(df_mgmt,
                        format = "html",
                        table.attr = "style='border: 1px solid black; border-collapse: collapse; width: 80%; margin-left: auto; margin-right: auto;'",
                        align = "c",
                        row.names = FALSE,
                        escape = FALSE)
            for (i in 1:(ncol(df_mgmt) - 1)) {
                tbl <- tbl %>% column_spec(i, border_right = "1px solid black")
            }
            # Apply italic to Genes column
            if (length(genes_col_idx) > 0) {
                tbl <- tbl %>% column_spec(genes_col_idx, italic = TRUE)
            }
            tbl %>%
                kable_styling(bootstrap_options = c("striped", "hover"),
                            full_width = FALSE,
                            position = "center")
        } else if (knitr::is_latex_output()) {
            tbl <- kable(df_mgmt,
                  format = "latex",
                  booktabs = TRUE,
                  align = "c",
                  row.names = FALSE) %>%
                kable_styling(latex_options = c("repeat_header", "striped", "hold_position", "scale_down"),
                            full_width = FALSE,
                            position = "center")
            # Apply italic to Genes column
            if (length(genes_col_idx) > 0) {
                tbl <- tbl %>% column_spec(genes_col_idx, italic = TRUE)
            }
            tbl
        } else {
            kable(df_mgmt, row.names = FALSE)
        }
    } else {
        # Handle empty file or header-only case
        if (knitr::is_html_output()) {
            cat("<p style='text-align: center; color: #666;'><em>No fusion event was detected in this sample.</em></p>")
        } else if (knitr::is_latex_output()) {
            cat("No fusion event was detected in this sample.")
        } else {
            cat("No fusion event was detected in this sample.")
        }
    }
}, error = function(e) {
    # Handle file reading errors with more detailed information
    error_msg <- paste("Error reading structure variant file:", e$message)
    cat(error_msg, "\n")
    
    if (knitr::is_html_output()) {
        cat("<p style='text-align: center; color: #666;'><em>No fusion event was detected in this sample.</em></p>")
    } else if (knitr::is_latex_output()) {
        cat("No fusion event was detected in this sample.")
    } else {
        cat("No fusion event was detected in this sample.")
    }
})
```

### Disclaimer 

This nanopore whole genome sequencing (nWGS) pipeline is an investigational research tool that has not undergone full clinical validation. Any clinical use or interpretation of its results is entirely at the discretion and responsibility of the treating physician

#### Report generated on `r Sys.time()`

