---
output:
  pdf_document:
    fig_height: 3
    latex_engine: pdflatex
header-includes:
   - \usepackage{helvet}
   - \renewcommand*\familydefault{\sfdefault}
   - \usepackage[T1]{fontenc}

geometry: margin=0.5in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(scales)
library(data.table)
library(rmarkdown)
library(dplyr)

votes <- fread(normalizePath(snakemake@input[["votes"]]))
dict <- fread(normalizePath(snakemake@input[["DICT_FILE"]]), check.names = T)

mc <- votes %>% 
 left_join(dict, by=c("class"="Methylation.Class.Name.Abbreviated"))

mcf <- mc %>%
 group_by(Methylation.Class.Family) %>%
 summarise(score_MCF = sum(score))
```

```{r, out.height= "50px", echo=FALSE, fig.align="right"}
knitr::include_graphics(file.path(snakemake@scriptdir, "header_PDFreport.pdf"))
```

## `r snakemake@params[["title"]]`

# Sample ID: `r snakemake@wildcards[["sample"]]`

## Quality control metrics

#### Barcode statistics

Demultiplexing statistics are shown for quality control only using a maximum of randomly subsampled 100,000 reads. All reads are considered for downstream analysis.

```{r comment='', echo=FALSE}
cat(readLines(normalizePath(snakemake@input[["demux"]])), sep = '\n')
```

#### Read statistics

```{r comment='', echo=FALSE}
cat(readLines(normalizePath(snakemake@input[["nanostat"]]), n=15), sep = '\n') # Read general summary (first 15 lines) of NanoStat statistics
```

Mean genome coverage is `r system(paste("awk 'END{print $4}' ", normalizePath(snakemake@input[["mosdepth"]]), sep=""), intern=T)`X.

```{r, out.height= "150px", echo=FALSE}
knitr::include_graphics(normalizePath(snakemake@input[["RL"]]))
knitr::include_graphics(normalizePath(snakemake@input[["GC"]]))
```

## Copy number profile

```{r, out.height = "180px", echo=FALSE}
knitr::include_graphics(normalizePath(snakemake@input[["CN"]]))
```

## Methylation-based classification

Methylation-based classification is based on **`r as.integer(votes$num_features[1])`** CpG sites (overlap of sites covered in this sample and the model).
At the methylation class (MC) level, the sample has been classified as **`r ifelse(mc$score[1] >= 0.2, mc$Methylation.Class.Name[1], "not classifiable (score < 0.2)")`**. 
This prediction has a confidence score of **`r round(mc$score[1], digits=3)`**.
At the methylation class **family** (MCF) level, the sample has been classified as **`r ifelse(max(mcf$score_MCF)>=0.2, mcf[which.max(mcf$score_MCF),]$Methylation.Class.Family, "not classifiable (score < 0.2)")`**.
The MCF prediction has a confidence score of **`r round(max(mcf$score_MCF), digits=3)`**.

Scores for the Top 5 entities on MC and MCF level are given below.
Vertical dashed lines indicate the recommended >0.2 cut-off for classification.

```{r, out.height = "150px", fig.align = 'center', echo=FALSE, warning=FALSE}
library(ggplot2)

ggplot(data=mc, aes(x=Methylation.Class.Name, y=score)) +
  geom_bar(stat="identity", fill="#1f78b4") +
  geom_hline(yintercept=0.2, linetype="dashed") + # indicate recommended cut-off
  coord_flip() +
  scale_x_discrete(limits=mc[order(mc$score,decreasing=T),]$Methylation.Class.Name[5:1]) +
  scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1), limits=c(0,1)) +
  ylab("score") + xlab("") +
  ggtitle("Methylation class (MC)") +
  theme_classic() + theme(aspect.ratio=1)
```

```{r, out.height = "150px", fig.align = 'center', echo=FALSE, warning=FALSE}
ggplot(mcf, aes(x=Methylation.Class.Family, y=score_MCF)) +
  geom_bar(stat="identity", fill="#4daf4a") +
  geom_hline(yintercept=0.2, linetype="dashed") + # indicate recommended cut-off
  coord_flip() +
  scale_x_discrete(limits=mcf[order(mcf$score_MCF,decreasing=T),]$Methylation.Class.Family[5:1]) +
  scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1), limits=c(0,1)) +
  ylab("score") + xlab("") +
  ggtitle("Methylation class family (MCF)") +
  theme_classic() + theme(aspect.ratio=1)
```

```{r dimensionality_reduction_plot, eval=snakemake@config[["dim_reduction_report"]], results='asis', echo=FALSE}
cat("## Dimensionality reduction plot\n\n")

cat("Dimensionality reduction plots are only intended for visual quality control, not classification, and must be interpreted with caution.\n\n")

knitr::include_graphics(normalizePath(snakemake@input[["tSNE"]]))
```

#### Disclaimer

Methylation-based classification using nanopore whole genome sequening is a research tool currently under development. It has not been clinically validated in sufficiently large cohorts. Interpretation and implementation of the results in a clinical setting is in the sole responsibility of the treating physician.

  
  

(Report generated on `r Sys.time()`. Pipeline version: `r system("git describe --tags", intern = T)`, `r system("git log --pretty=format:'%h' -n 1", intern=T)`)
